---
title: "Statistical Concepts"
author: "Yazan Dabbas"
date: "12/12/2023"
format: 
    html:
      code-line-numbers: true
      df-print: paged
---
**Load the Data **

------------------------------------------------------------------------
```{r}

# Load the data
df <- readRDS("C:/Master Degree/Core Qualification_Compulsory Courses/Business/Causal_Data_Science/Data/Causal_Data_Science_Data/random_vars.rds")

# Be familiar with the data
head(df)
```
**Get an Overview of the Data **

------------------------------------------------------------------------
```{r}

View(df)

```


```{r}

# Store columns in variables
age <- df$age
income <- df$income

# 1) 
```
**a) Calculate the Expected Value **

------------------------------------------------------------------------
```{r}

expected_age <- sum(age) / length(age)
expected_income <- sum(income) / length(income)

cat("Expected age is ", expected_age, "\n")
cat("Expected income is ", expected_income, "\n")
```
**b)  Calculate the Variance **

------------------------------------------------------------------------
```{r}
mean_age <- mean(age)
mean_income <- mean(income)

variance_age <- sum((age - mean_age)^2) / (length(age) - 1)
variance_income <- sum((income - mean_income)^2) / (length(income) - 1)

cat("Age Variance is ", variance_age, "\n")
cat("Income Variance is ", variance_income, "\n")
```
**c)  Calculate the Standard Deviation **

------------------------------------------------------------------------
```{r}
SD_age <- sqrt(variance_age)
SD_income <- sqrt(variance_income)
cat("SD of Age is ", SD_age, "\n")
cat("SD of Income is ", SD_income, "\n")

```
**2) Standard Deviation Comparison**

------------------------------------------------------------------------

::: {.callout-important appearance="minimal"} 

-   **When the variables are measured in the same units and have similar scales, comparing standard deviations directly is most useful.** 

-   **In the instance of age and income, the variables have distinct units and scales. As a result, directly comparing their standard deviations may not yield relevant insights.**

:::

**3) The Relationship Between Covariance and Correlation**

------------------------------------------------------------------------
```{r}
covariance_value <- sum((age - mean_age) * (income - mean_income)) /length(age)
cat("Covariance Value:", covariance_value, "\n")


correlation_value <- covariance_value / (SD_age * SD_income)
cat("Correlation Value:", correlation_value, "\n")
```

**4) Interpretation regarding the Relationship Between Covariance and Correlation**

------------------------------------------------------------------------

::: {.callout-important appearance="minimal"} 

-   **The standardized scale of the correlation coefficient makes comparison easier. The fact that it goes from `-1` to `1` gives for a clear grasp of the relationship's strength and direction.**

:::
 
 **5) Compute the Conditional Expected Value for:**
 
------------------------------------------------------------------------
```{r}
 # 1. E[income|age<=18]

subset_1 <-  df %>% filter(age <= 18)
conditional_expected_value1 <- mean(subset_1$income)
cat("Conditional Expected Value of Income for age <= 18:", conditional_expected_value1, "\n")

# 2. E[income|age<=[18,65)]

subset_2 <- df %>% filter(age >= 18 & age < 65)
conditional_expected_value2 <- mean(subset_2$income)
cat("Conditional Expected Value of Income for age in [18, 65):", conditional_expected_value2, "\n")

#3. E[income|age>=65]

subset_3 <- df %>% filter(age >= 65)
conditional_expected_value3 <- mean(subset_3$income)
cat("Conditional Expected Value of Income for age >= 65:", conditional_expected_value3, "\n")


```