---
title: "Difference-in-Differences"
author: "Yazan Dabbas"
date: "12/22/2023"
format: 
    html:
      code-line-numbers: true
      df-print: paged
---

```{r}
# Load necessary libraries
library(dplyr)
library(readr)
library(lmtest)
```

```{r}
# Load the data
data <- readRDS('C:/Master Degree/Core Qualification_Compulsory Courses/Business/Causal_Data_Science/Data/Causal_Data_Science_Data/hospdd.rds')

# Be familiar with the data
head(data)

```



#### **1. Manually compute the mean satisfaction for treated and control hospitals before and after the treatment**

----
```{r}
# The month after that they introduced the new admission 
threshold_month <- 3.0

# The hospital-ID, after which the others are not treated
threshold_hospital <- 18

# Convert Month and hospital columns into numeric
data$month <- as.numeric(data$month)
data$hospital <- as.numeric(data$hospital)

# Split the data into treated and controlled groups
treated_group <- data %>%
  filter(hospital <= 18 )

control_group <- data %>%
  filter(hospital > 18 )


# Mean Difference between treatment and control group BEFORE treatment
before_control_mean <- control_group %>% 
  filter(month <= threshold_month) %>%
  summarise(mean_satisfaction = mean(satis)) %>%
  pull(mean_satisfaction)

before_treatment_mean  <- treated_group %>% 
  filter( month <= threshold_month) %>%
  summarise(mean_satisfaction = mean(satis)) %>%
  pull(mean_satisfaction)


mean_diff_before <- before_treatment_mean - before_control_mean


# Mean Difference between treatment and control group AFTER treatment
after_control_mean <- control_group %>% 
  filter(month > threshold_month) %>%
  summarise(mean_satisfaction = mean(satis)) %>%
  pull(mean_satisfaction)

after_treatment_mean <- treated_group %>% 
  filter( month > threshold_month) %>%
  summarise(mean_satisfaction = mean(satis)) %>%
  pull(mean_satisfaction)


mean_diff_after <- after_treatment_mean - after_control_mean


# Difference-in-differences

mean_diff_diff <- mean_diff_after - mean_diff_before


```
#### **2. Use linear regression to compute the estimate with group and time fixed effects**

----
```{r}

# Fit the linear regression model
model <- lm(satis ~ procedure * as.factor(month) + as.factor(hospital) , data)

# Print the results
summary(model)

```

::: {.callout-important appearance="minimal"}

-   **When you use `month + hospital` in regression, R treats these variables as numeric, and each level gets its own coefficient. This implies a linear trend for months and a linear effect for each hospital. This might not be appropriate for months and hospitals, which are better modeled as categorical factors.**

-   **On the other hand, `as.factor(month) + as.factor(hospital)` treats both months and hospitals as factors, assigning a separate intercept for each level. This is more appropriate when dealing with categorical variables like months and hospitals because it allows for different intercepts for each level rather than assuming a linear relationship.**'

-   **Using `as.factor(month) + as.factor(hospital)` is more suitable for this analysis, as it captures the categorical nature of months and hospitals, allowing for separate effects for each level**

:::