---
title: "Instrumental Variables"
author: "Yazan Dabbas"
date: "12/25/2023"
format: 
    html:
      code-line-numbers: true
      df-print: paged
---

```{r}

# Load required libraries
library(tidyverse)
library(dagitty)
library(ggdag)
library(estimatr)
library(AER)
```

**First Question: **

------------------------------------------------------------------------

```{r}
# Define the DAG
dag <- dagify(
  spentTime ~ featureUsed,
  spentTime ~ Unobserved,
  featureUsed ~ Unobserved,
  featureUsed ~ encourgement,
  exposure = "featureUsed",
  latent = "Unobserved",
  outcome = "spentTime",
  coords = list(x = c(Unobserved = 1, featureUsed = 0, spentTime = 2, encourgement = -1),
                y = c(Unobserved = 1, featureUsed = 0, spentTime = 0, encourgement = 0)),
  labels = c(
    "spentTime" = "Time Spent on the App",
    "featureUsed" = "The new feature is used",
    "encourgement" = "User encourgement to use feature",
    "Unobserved" = "Unobserved variables"
  )
)
# Plot DAG
ggdag(dag, text = FALSE, use_labels = "label")


```

```{r}

#Load the data
df <- readRDS('C:/Master Degree/Core Qualification_Compulsory Courses/Business/Causal_Data_Science/Data/Causal_Data_Science_Data/rand_enc.rds')


# Be familiar with the data

head(df)

```

**Second Question: **

------------------------------------------------------------------------

```{r}
#| warning: false


# Compute Naive Estimate
naive_estimate <- lm(time_spent ~ used_ftr, data = df)

summary(naive_estimate)
```

**Third Question: **

------------------------------------------------------------------------
```{r}
# Check the correlation matrix
cor(df) %>% round(2)

```
  
::: {.callout-important appearance="minimal"} 

-   **Since the naive estimate (10.82269) is greater than the IV robust estimate using rand_enc as an instrument (9.738175), we would consider the naive estimate to have an upward bias.**

-   **This implies that the naive estimate overestimates the effect of `used_ftr` on `time_spent`.**
  
-   **A strong correlation between `used_ftr` and `time_spent` can be seen.**

-   **Assuming `rand_enc` as an Instrumental variable would make sense, because it has a weak correlation with the outcome `(time_spent)` and a stronger one with the treatment `(used_ftr)`.**

-   **While the correlation between Instrumental variable and outcome is not  zero (maybe due to noise), this correlation is relatively low.**

:::


**Fourth Question: **

------------------------------------------------------------------------
```{r}

# Instrumental Variable Estimation using 2SLS with rand_enc and robust standard errors

model_iv_robust <- iv_robust(time_spent ~ used_ftr | rand_enc, data = df)
summary(model_iv_robust)


## Hansen J test

# Extract residuals and fitted values from the model
residuals_iv <- residuals(model_iv_robust)
fitted_values_iv <- fitted(model_iv_robust)

# Perform Hansen J test
hansen_test_stat <- sum(residuals_iv * fitted_values_iv)
p_value_hansen <- 1 - pchisq(hansen_test_stat, df = 1)

# Display results
cat("Hansen J Test Statistic:", hansen_test_stat, "\n")
cat("P-value:", p_value_hansen, "\n")
```

::: {.callout-important appearance="minimal"}

-   **Hansen J test with a test statistic close to 0 and a p-value close to 1 indicates that the instrument used in the model is not violating the over-identifying restrictions.**

-   **In other words, the instrument is valid for the model, and there is no evidence to suggest that the instrument is endogenous or correlated with the error term.** 
  
:::

```{r}
cat("Naive Estimate:", coef(naive_estimate)['used_ftr'], "\n")
cat("IV Robust Estimate (rand_enc):", model_iv_robust$coefficients['used_ftr'], "\n")
```

::: {.callout-important appearance="minimal"}

-   **Since the naive estimate (10.82269) is greater than the IV robust estimate using rand_enc as an instrument (9.738175), we would consider the naive estimate to have an upward bias.**

-   **This implies that the naive estimate overestimates the effect of used_ftr on time_spent.**
  
:::
