---
title: "Regression Discontinuity"
author: "Yazan Dabbas"
date: "12/26/2023"
format: 
    html:
      code-line-numbers: true
      df-print: paged

---
----
```{r}

# Load required libraries
library(dplyr)
library(ggplot2)


# Read data for the current campaign
df <- readRDS('C:/Master Degree/Core Qualification_Compulsory Courses/Business/Causal_Data_Science/Data/Causal_Data_Science_Data/coupon.rds')

```

#### **Regression Discontinuity Design Sensitivity Analysis**

----
```{r}
# Define cut-off
c0 <- 60

# Bandwidths
bw_orignal <- c0 + c(-5,5)
bw_half <- c0 + c(-5,5) / 2
bw_double <- c0 + c(-5,5) * 2

```

 **Function to run the regression discontinuity design analysis**
 
----
```{r}



run_rdd_analysis <- function(df, bw) {
  
  df_bw_below <- df %>% filter(days_since_last >= bw[1] & days_since_last < c0)
  df_bw_above <- df %>% filter(days_since_last >= c0 & days_since_last <= bw[2])
  df_bw <- bind_rows(df_bw_above, df_bw_below)
  
  lm_bw <- lm(purchase_after ~ days_since_last_centered + coupon, df_bw)
  
  
  model_bw_below <- lm(purchase_after ~ days_since_last, df_bw_below)
  model_bw_above <- lm(purchase_after ~ days_since_last, df_bw_above)
  
  y0 <- predict(model_bw_below, tibble(days_since_last = c0))
  y1 <- predict(model_bw_above, tibble(days_since_last = c0))
  
  late <- y1 - y0
  return(list(LATE = late, Summary = summary(lm_bw)))
}

```

 **Use the Function for Different Bandwidths**
 
----
```{r}
# Run the analysis with original bandwidth
LATE_original <- run_rdd_analysis(df, bw_orignal)

# Run the analysis with half the bandwidth
LATE_half_bandwidth <- run_rdd_analysis(df, bw_half)

# Run the analysis with double the bandwidth
LATE_double_bandwidth <- run_rdd_analysis(df, bw_double)

```
----
1. **Original Bandwidth**
----

The LATE and the Summary of Using the Original Bandwidth
```{r}
#| code-fold: true

# Print the results for the original bandwidth
cat("LATE:", LATE_original$LATE, "\n")
print(LATE_original$Summary)
```
----
2.  **Half the Bandwidth**
----

The LATE and the Summary of Using Half the Bandwidth
```{r}
#| code-fold: true

# Print the results for half the bandwidth
cat("LATE:", LATE_half_bandwidth$LATE, "\n")
print(LATE_half_bandwidth$Summary)
```
----
3.  **Double the Bandwidth**
----

The LATE and the Summary of Using Double the Bandwidth
```{r}
#| code-fold: true


# Print the results for double the bandwidth
cat("LATE:", LATE_double_bandwidth$LATE, "\n")
print(LATE_double_bandwidth$Summary)
```

::: {.callout-important appearance="minimal"}
-   **There is no change in the three bandwidth choices on the statistical effect (positive effect of coupon variable on purchase_after.**

-   **The estimated LATE is slightly lower for the half bandwidth than the original bandwidth, suggesting a more conservative estimate.**

-   **The estimated LATE is higher for the double bandwidth than the original bandwidth, suggesting a potentially broader impact on individuals farther from the cutoff point.**

-   **Bandwidth choice influences treatment `(coupuonTRUE)` effect estimation.**

-   **Coefficients vary with bandwidth changes**
:::

### **Different Past Campaign**

----
```{r}
#| warning: false

# Read data for the different past campaign

df_shipping <- readRDS('C:/Master Degree/Core Qualification_Compulsory Courses/Business/Causal_Data_Science/Data/Causal_Data_Science_Data/shipping.rds')

# Manipulation Testing Using Local Polynomial Density Estimation
library(rddensity) 
rddd <- rddensity(df_shipping$purchase_amount, c = 30) 

summary(rddd)

```

::: {.callout-important appearance="minimal"}
-   **The manipulation tests indicate significant evidence of manipulation around the cut-off point (c=30).**

-   **P-values of the robust T-statistic are close to zero, suggesting that there are systematic changes in the observed density of the `purchase_amount` variable near the cut-off.**

-   **The order of estimation and bias is also different on each side, suggesting a lack of smoothness or continuity.**

-   **The p-values of binomial tests provide further evidence of non-random behavior around the cut-off.**

-   **The `purchase_amount` variable, based on the results of manipulation testing, may not be appropriate as a running variable for an RDD with a cut-off at 30€.**
:::
----
**Plot to confirm that `purchase_amount` could not be used as a running variable at 30**
----
```{r}

ggplot(df_shipping, aes(x = purchase_amount)) + geom_histogram(binwidth = 5, fill = "blue", color = "white") + geom_vline(xintercept = 30, color = "red", linetype = "dashed") + xlab("Purchase Amount (€)") + ylab("Number of Purchases") + theme_minimal()

```